Article data analysis 2

### 2019-03-27 notebooks/article_notebook2.Rmd
minor changes

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
# WARNING: This will install packages on your system if not present in library path!
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, ape, cluster, tibble, Biostrings, seqinr, stringr, scales, reprex)
```

@@

# Genes of interest. These genes are partially matched to 
# the generated gene names from the "fix_gene_names" function
ac_genes <- c("qnr","oqx","qep","aac6")
mut_genes <- c("gyr","par","marR","marA",
               "sox","rpoB","rob","gad")

@@

  # Left join to match with ref_name column
  # and strip ref column for file suffix
  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))

@@

@@

## Specific mutations in non-QRDR intrinsic genes
These mutations are the most prominent in each gene where mutations were identified (n > 2). The percentage is based on how many of the total isolates with mutations in the respective gene had the specific mutation in question.
```{r, message=FALSE, warning=FALSE}
mut_filtered %>%
  filter(gene %not_in% c("gyrA","gyrB","parC","parE")) %>%
  select(id, gene, mut) %>%
  mutate(entries = sapply(strsplit(.$mut, ","),
                          FUN = function(x) {length(x)})) %>%
  separate(mut, into = paste("v", 1:max(.$entries), sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:max(.$entries), sep = "_")) %>%
  mutate(value = str_trim(value)) %>%
  select(id, gene, value) %>%
  filter(value != 0,
         is.na(value) == FALSE) %>%
  group_by(gene, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  filter(n > 2) %>%
  mutate(percent = case_when(
    gene == "gadX" ~ 
      round(n / length(which(mut_report[, "gadX"] == 1))*100, 1),
    gene == "marA" ~ 
      round(n / length(which(mut_report[, "marA"] == 1))*100, 1),
    gene == "marR" ~ 
      round(n / length(which(mut_report[, "marR"] == 1))*100, 1),
    gene == "rpoB" ~ 
      round(n / length(which(mut_report[, "rpoB"] == 1))*100, 1),
    gene == "soxS" ~ 
      round(n / length(which(mut_report[, "soxS"] == 1))*100, 1))) %>%
  arrange(gene, desc(percent)) %>%
  dplyr::rename("Gene" = gene,
                "Mutation" = value,
                "Percent" = percent)

@@

## Total combinations of resistance mechanisms
This table presents all combinations of resistance mechanisms included in this study, and their respective MIC.
```{r}
mut_report %>%
  select(-c(nt_change, gyrA, parC, parE)) %>%
  left_join(mut_quant[c("gyrA","parC","parE","id")], by = "id") %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  # summarise qnr-genes
  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  arrange(desc(CIP))
```

@@

## Double-check ARIBA results
This chunk double-checks the mutations found in _gyrA_, _parC_ and _parE_ against an internal mutation analysis script. The sequences are taken from Prokka, and reference wild-type versions of the genes are used to identify mismatches with the sample sequences.
```{r, warning=FALSE, message=FALSE}
# Functions
mismatches <- function(query, ref) {
  pairwiseAlignment(ref, query, substitutionMatrix = "BLOSUM50",
                    gapOpening = 3, gapExtension = 1) %>%
    mismatchTable() %>%
    mutate(ID=names(query),
           Pos=PatternStart,
           Reference_AA=as.character(PatternSubstring),
           Sample_AA=as.character(SubjectSubstring)) %>%
    select(ID, Reference_AA, Sample_AA, Pos) %>%
    filter(Sample_AA != "X")
}

@@

mlst_mechanisms_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,14:20]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4,nt_change)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mlst_data[c("id","ST")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, ST), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,ST)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit() %>%
  group_by(ST, gene, value) %>%
  dplyr::count() %>%
  filter(ST %in% no_of_mlst$ST)

@@

## Percent resistance mechanisms in total
```{r}
total_mechanisms <- mut_report %>%
  select(-nt_change) %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, result, -id) %>%
  group_by(gene, result) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(result = ifelse(result == 1, "Present", "Absent")) %>%
  spread(result, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round((Present/Total)*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1),
         type = ifelse(gene %in% c("qnr","qepA4"), "Acquired","Intrinsic"))

@@

mut_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    mutate(species = factor(species,
                            levels = c("Broiler","Pig","Wild bird","Red fox"))) %>%
    ggplot(aes(reorder(key, -Percent), Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_fill_manual(values = palette) +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

@@

acquired_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, qnr_genes) %>%
    group_by(key, value, species) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(0.9)) +
    scale_fill_manual(values = palette) +
    labs(y = "Percent (%) of isolates",
         fill = NULL) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.justification = c(0, 1),
      legend.position = c(0.82, 0.97)
    )
```

@@

corr_matrices <- lapply(total_df, function(x) create_corr_matrix(x)) %>%
  bind_rows(.id = "species") %>%
  mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
           gene2 = gsub("(.*?)_.+", "\\1", var2),
           var1 = gsub(".+_(.*?)", "\\1", var1),
           var2 = gsub(".+_(.*?)", "\\1", var2),
           type1 = case_when(gene1 == "gyrA" ~ 1,
                          gene1 == "gyrB" ~ 2,
                          gene1 == "parC" ~ 3,
                          gene1 == "parE" ~ 4,
                          gene1 == "marR" ~ 5,
                          gene1 == "marA" ~ 6,
                          gene1 == "rpoB" ~ 7,
                          gene1 == "soxS" ~ 8,
                          gene1 == "robA" ~ 9,
                          gene1 == "gadX" ~ 10,
                          gene1 == "qnrA1" ~ 11,
                          gene1 == "qnrB19" ~ 12,
                          gene1 == "qnrB6" ~ 13,
                          gene1 == "qnrB60" ~ 14,
                          gene1 == "qnrS1" ~ 15,
                          gene1 == "qnrS2" ~ 16,
                          gene1 == "qnrS4" ~ 17,
                          gene1 == "qnr" ~ 18,
                          gene1 == "qepA4" ~ 19),
           type2 = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "marR" ~ 5,
                          gene2 == "marA" ~ 6,
                          gene2 == "rpoB" ~ 7,
                          gene2 == "soxS" ~ 8,
                          gene2 == "robA" ~ 9,
                          gene2 == "gadX" ~ 10,
                          gene2 == "qnrA1" ~ 11,
                          gene2 == "qnrB19" ~ 12,
                          gene2 == "qnrB6" ~ 13,
                          gene2 == "qnrB60" ~ 14,
                          gene2 == "qnrS1" ~ 15,
                          gene2 == "qnrS2" ~ 16,
                          gene2 == "qnrS4" ~ 17,
                          gene2 == "qnr" ~ 18,
                          gene2 == "qepA4" ~ 19))

@@

## MIC-plots
```{r}
total_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  dplyr::select(-nt_change) %>%
  mutate_at(vars(-id),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  select(-ref) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(present,CIP) %>%
  dplyr::count() %>%
  ungroup() %>%
  rowwise() %>%
  mutate(present = ifelse(present == "", "None", present),
         total = 280,
         percent = round(n/total * 100, 1),
         lwr = round(get_binCI(n, total)[1], 1),
         upr = round(get_binCI(n, total)[2], 1)) %>%
  ungroup() %>%
  mutate(entries = sapply(strsplit(.$present, ","),
                          FUN = function(x) {length(x)}))

@@

dot_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  select(-nt_change) %>%
  mutate_at(vars(-id),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,13:19]) > 0, 1, 0)) %>%
  select(-c(qnrA1,qnrB19,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,CIP)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit() %>%
  mutate(entries = sapply(strsplit(.$present, ","),
                          FUN = function(x) {length(x)}))

dot_plot <- ggplot(dot_report, aes(reorder(present, entries), gene, fill = gene))+
  geom_point(size = 2)+
  scale_y_discrete(limits = c("gyrA","parC","parE","gadX","marA","marR","rpoB","soxS","qepA4","qnr"))+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        axis.ticks.x = element_blank())

>>>>

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(viridis)
library(purrr)
library(tibble)
library(stringr)
library(Biostrings)
library(seqinr)
library(scales)
library(gridExtra)
```

@@

# Genes of interest. These genes are partially matched to 
# the generated gene names from the "fix_gene_names" function
ac_genes <- c("qnr","oqx","qep","aac6")
mut_genes <- c("gyr","par","marR","marA",
               "sox","rpoB","rob")

@@

  # Left join to match with ref_name column
  # and strip ref column for file suffix
  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref)) %>%
    mutate(gene_names = ifelse(gene_names == "soxS", "soxR", gene_names))

@@

## Percent mutations in each intrinsic gene per animal species
```{r}
mut_report %>%
  select(-nt_change) %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  gather(gene, value, -c(id, species)) %>%
  group_by(gene, species, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total * 100,1)) %>%
  arrange(gene)
```

@@

## Specific mutations in non-QRDR intrinsic genes
These mutations are the most prominent in each gene where mutations were identified (n > 2). The percentage is based on how many of the total isolates with mutations in the respective gene had the specific mutation in question.
```{r, message=FALSE, warning=FALSE}
mut_filtered %>%
  filter(gene %not_in% c("gyrA","gyrB","parC","parE")) %>%
  select(id, gene, mut) %>%
  mutate(entries = sapply(strsplit(.$mut, ","),
                          FUN = function(x) {length(x)})) %>%
  separate(mut, into = paste("v", 1:max(.$entries), sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:max(.$entries), sep = "_")) %>%
  mutate(value = str_trim(value)) %>%
  select(id, gene, value) %>%
  filter(value != 0,
         is.na(value) == FALSE) %>%
  group_by(gene, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  filter(n > 2) %>%
  mutate(Total = case_when(
    gene == "marA" ~
      length(which(mut_report[, "marA"] == 1)),
    gene == "marR" ~
      length(which(mut_report[, "marR"] == 1)),
    gene == "rpoB" ~
      length(which(mut_report[, "rpoB"] == 1)),
    gene == "soxR" ~
      length(which(mut_report[, "soxR"] == 1))),
         percent = case_when(
    gene == "marA" ~ 
      round(n / Total*100, 1),
    gene == "marR" ~ 
      round(n / Total*100, 1),
    gene == "rpoB" ~ 
      round(n / Total*100, 1),
    gene == "soxR" ~ 
      round(n / Total*100, 1))) %>%
  arrange(gene, desc(percent)) %>%
  dplyr::rename("Gene" = gene,
                "Substitution" = value,
                "Percent" = percent)

@@

## Total combinations of resistance mechanisms
This table presents all combinations of resistance mechanisms included in this study, and their respective MIC.
```{r}
mut_report %>%
  select(-c(nt_change, gyrA, parC, parE)) %>%
  left_join(mut_quant[c("gyrA","parC","parE","id")], by = "id") %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  # summarise qnr-genes
  mutate(qnr = ifelse(rowSums(.[,12:18]) == 0, 0, 1)) %>%
  left_join(mic_data[c("id","CIP","NAL")], by = "id") %>%
  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  arrange(desc(CIP))
```

@@

```{r, include=FALSE}
# Functions
mismatches <- function(query, ref) {
  pairwiseAlignment(ref, query, substitutionMatrix = "BLOSUM50",
                    gapOpening = 3, gapExtension = 1) %>%
    mismatchTable() %>%
    mutate(ID=names(query),
           Pos=PatternStart,
           Reference_AA=as.character(PatternSubstring),
           Sample_AA=as.character(SubjectSubstring)) %>%
    select(ID, Reference_AA, Sample_AA, Pos) %>%
    filter(Sample_AA != "X")
}

@@

mlst_mechanisms_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,13:19]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4,nt_change)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mlst_data[c("id","ST")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, ST), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,ST)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit() %>%
  group_by(ST, gene, value) %>%
  dplyr::count() %>%
  filter(ST %in% no_of_mlst$ST)

@@

## Percent resistance mechanisms in total
```{r}
total_mechanisms <- mut_report %>%
  select(-nt_change) %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,13:18]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, result, -id) %>%
  group_by(gene, result) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(result = ifelse(result == 1, "Present", "Absent")) %>%
  spread(result, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round((Present/Total)*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1),
         type = ifelse(gene %in% c("qnr","qepA4"), "Acquired","Intrinsic"))

@@

mut_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    mutate(species = factor(species,
                            levels = c("Broiler","Pig","Wild bird","Red fox")),
           key = paste0(toupper(substr(key, 1, 1)),
                        substr(key, 2, nchar(key))),
           group = if_else(key %in% c("GyrA","GyrB","ParC","ParE"), 1, 2)) %>%
  ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_fill_manual(values = palette) +
    scale_x_discrete(labels = c("GyrA","GyrB","ParC","ParE",
                                "MarA","MarR","SoxR","RpoB",
                                "RobA")) +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

@@

acquired_report %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, qnr_genes) %>%
    group_by(key, value, species) %>%
    dplyr::count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(0.9)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(0.9)
    ) +
    scale_fill_manual(values = palette) +
    labs(y = "Percent (%) of isolates",
         fill = NULL) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 12),
      axis.text.x = element_text(face = "italic"),
      axis.title.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.justification = c(0, 1),
      legend.position = c(0.82, 0.97)
      )
```

@@

corr_matrices <- lapply(total_df, function(x) create_corr_matrix(x)) %>%
  bind_rows(.id = "species") %>%
  mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
           gene2 = gsub("(.*?)_.+", "\\1", var2),
           var1 = gsub(".+_(.*?)", "\\1", var1),
           var2 = gsub(".+_(.*?)", "\\1", var2),
           type1 = case_when(gene1 == "gyrA" ~ 1,
                          gene1 == "gyrB" ~ 2,
                          gene1 == "parC" ~ 3,
                          gene1 == "parE" ~ 4,
                          gene1 == "marR" ~ 5,
                          gene1 == "marA" ~ 6,
                          gene1 == "rpoB" ~ 7,
                          gene1 == "soxR" ~ 8,
                          gene1 == "robA" ~ 9,
                          gene1 == "qnrA1" ~ 10,
                          gene1 == "qnrB19" ~ 11,
                          gene1 == "qnrB6" ~ 12,
                          gene1 == "qnrB60" ~ 13,
                          gene1 == "qnrS1" ~ 14,
                          gene1 == "qnrS2" ~ 15,
                          gene1 == "qnrS4" ~ 16,
                          gene1 == "qnr" ~ 17,
                          gene1 == "qepA4" ~ 18),
           type2 = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "marR" ~ 5,
                          gene2 == "marA" ~ 6,
                          gene2 == "rpoB" ~ 7,
                          gene2 == "soxR" ~ 8,
                          gene2 == "robA" ~ 9,
                          gene2 == "qnrA1" ~ 10,
                          gene2 == "qnrB19" ~ 11,
                          gene2 == "qnrB6" ~ 12,
                          gene2 == "qnrB60" ~ 13,
                          gene2 == "qnrS1" ~ 14,
                          gene2 == "qnrS2" ~ 15,
                          gene2 == "qnrS4" ~ 16,
                          gene2 == "qnr" ~ 17,
                          gene2 == "qepA4" ~ 18))

@@

## Mechanisms per animal species
```{r, fig.height=10, fig.width=15, warning=FALSE, message=FALSE}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    unlist() %>% unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

total_df <- acquired_report %>%
  left_join(mut_quant[c("gyrA","gyrB","parC","parE","id")], by = "id") %>%
  left_join(mut_report[c("id","marR","marA","rpoB","soxR","robA")], by = "id") %>%
  left_join(isolate_data[c("id", "species")])

names(total_df) <- c("id","qepA4","qnrA1","qnrB19",
                     "qnrB6","qnrB60","qnrS1","qnrS2",
                     "qnrS4","gyrA","gyrB","parC",
                     "parE","MarR","MarA","RpoB",
                     "SoxR","RobA","species")

id <- total_df$id

cols_to_add <- sapply(function(i) {AddCol(df = total_df, col_name = col_proc[i])},
                      X = seq_along(col_proc)) %>%
  bind_cols() %>%
  cbind(id) %>%
  gather(key, value, -id) %>%
  filter(grepl("(.*?)_0", key) == FALSE) %>%
  spread(key, value)

total_df2 <- total_df %>%
  select(-c(gyrA, gyrB, parC, parE)) %>%
  left_join(cols_to_add, by = "id") %>%
  select(id, species, everything()) %>%
  mutate_at(vars(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,4:10]) == 0, 0, 1)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-id) %>%
  mutate_at(vars(-species, CIP),
            funs(as.numeric(as.character(.)))) %>%
  group_by(species, CIP) %>%
  summarise_all(funs(sum, n = n())) %>%
  dplyr::rename("n" = MarR_n) %>%
  select(-contains("_n")) %>%
  gather(gene, value, -c(species, CIP, n)) %>%
  mutate(gene2 = sub("(.*?)_(.*?)_sum", "\\1", gene),
         gene2 = sub("(.*?)_sum", "\\1", gene2),
         gene = sub("(.*?)_(.*?)_sum", "\\2", gene),
         gene = sub("(.*?)_sum", "\\1", gene),
         type = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "MarR" ~ 5,
                          gene2 == "MarA" ~ 6,
                          gene2 == "RpoB" ~ 7,
                          gene2 == "SoxR" ~ 8,
                          gene2 == "RobA" ~ 9,
                          gene2 == "qnr" ~ 10,
                          gene2 == "qepA4" ~ 11),
         perc = value/n*100)

p1 <- ggplot(total_df2, aes(reorder(gene, type), factor(CIP), fill = perc))+
  geom_tile()+
  scale_fill_viridis(option = "A",
                     begin = 0.1,
                     guide = guide_colorbar(barheight = 10,
                                            barwidth = 0.8,
                                            label.position = "left",
                                            label.hjust = 0.5,
                                            label.vjust = 0.3,
                                            ticks = TRUE))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        strip.background = element_rect(colour = "black", fill = "grey70"),
        strip.text = element_text(size = 12),
        legend.position = "left",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        plot.margin = unit(c(0,0,0,0), "cm"))+
  labs(y = "CIP MIC")+
  facet_wrap(~species, ncol = 1)+
  coord_fixed()

nplot_data <- total_df %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(species, CIP) %>%
  dplyr::count() %>%
  ungroup()

n_names <- c("Broiler" = "n",
             "Pig" = "n",
             "Red fox" = "n",
             "Wild bird" = "n")

p2 <- ggplot(nplot_data, aes(factor(CIP), n))+
  geom_col()+
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.ticks.y = element_blank(),
        strip.text = element_text(size = 12, face = "italic"),
        strip.background = element_rect(colour = "black", fill = "grey70"),
        plot.margin = unit(c(0,0,-10,0), "cm"))+
  facet_wrap(~species, ncol = 1, labeller = as_labeller(n_names))+
  coord_flip()

p1g <- ggplotGrob(p1)
p2g <- ggplotGrob(p2)

g <- cbind(p1g, p2g, size = "first")

grid.arrange(g)
```

## MIC-plots
```{r}
total_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  dplyr::select(-nt_change) %>%
  mutate_at(vars(-id),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,12:18]) > 0, 1, 0)) %>%
  select(-c(qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  select(-ref) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(present,CIP) %>%
  dplyr::count() %>%
  ungroup() %>%
  rowwise() %>%
  mutate(present = ifelse(present == "", "None", present),
         total = 280,
         percent = round(n/total * 100, 1),
         lwr = round(get_binCI(n, total)[1], 1),
         upr = round(get_binCI(n, total)[2], 1)) %>%
  ungroup() %>%
  mutate(entries = sapply(strsplit(.$present, ","),
                          FUN = function(x) {length(x)}))

@@

dot_report <- mut_report %>%
  left_join(acquired_report, by = "id") %>%
  select(-nt_change) %>%
  mutate_at(vars(-id),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,12:18]) > 0, 1, 0)) %>%
  select(-c(qnrA1,qnrB19,qnrS1,qnrS2,qnrS4)) %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,CIP)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit() %>%
  mutate(entries = sapply(strsplit(.$present, ","),
                          FUN = function(x) {length(x)}))

dot_plot <- ggplot(dot_report, aes(reorder(present, entries), gene, fill = gene))+
  geom_point(size = 2)+
  scale_y_discrete(limits = c("gyrA","parC","parE","marA","marR","rpoB","soxR","qepA4","qnr"))+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        axis.ticks.x = element_blank())
